Local Spectral Clustering of Density Upper Level Sets

### 2019-04-29 arxiv/arxiv_local_density_clustering_proofs.tex
added arxiv version of paper

\documentclass{article}

% \usepackage{icml2019}
% \usepackage[accepted]{icml2019}

\renewcommand{\thesection}{\Alph{section}}
\renewcommand{\theequation}{A.\arabic{equation}}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{xr-hyper}
% \usepackage{hyperref}
\usepackage{bm}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{natbib}

\externaldocument{icml_local_density_clustering}

@@

We will fix $\Aset \subset \Rd$ to be an arbitrary set. To simplify expressions, for the $\sigma$-expansion $\Asig$, we will write the set difference between $\Asig$ and the $(\sigma + r)$-expansion $\Aset_{\sigma + r}$ as 
\begin{equation*}
\Asigr := \set{x: 0 < \dist(x, \Asig) \leq r},
\end{equation*}
where as a reminder $\dist(x, \Aset) = \min_{x' \in \Aset} \norm{x - x'}$.

For notational ease, we write
\begin{align*}
\cut_{n,r} = \cut(\Csig[\Xbf]; G_{n,r}), ~ \mu_K = \mathbb{E}(\cut_{n,r}), ~ p_K = \frac{\mu_K}{{n \choose 2}} \\
\vol_{n,r} = \vol(\Csig[\Xbf]; G_{n,r}), ~ \mu_V = \mathbb{E}(\vol_{n,r}), ~ p_V = \frac{\mu_V}{{n \choose 2}}
\end{align*}
for the random variable, mean, and probability of cut size and volume, respectively.

We state Lemma \ref{lem: expansion_sets} without proof, as it is trivial. We formally include it mainly to comment on its (potential) suboptimality; for sets $\Aset$ with diameter much larger than $\sigma$, the volume estimate of Lemma \ref{lem: expansion_sets} will be quite poor.

\begin{lemma}
	\label{lem: expansion_sets}
	For any $\delta > 0$ and $x \in \Aset_{\sigma}$,
	\begin{equation*}
	\sigma B \subset \Asig, ~~\mathrm{and~ }\nu(\Asig + \delta B) \leq \nu\left(\left[1 + \frac{\delta}{\sigma}\right]\Asig\right) = \left(1 + \frac{\delta}{\sigma}\right)^d \nu(\Asig)
	\end{equation*}
	where $\Asig = \Aset + \sigma B$ is the $\sigma$-expansion of $\Aset$.
\end{lemma}

We will need to carefully control the volume of the expansion set using the above estimate; Lemma \ref{lem: Taylor_series} serves this purpose.
\begin{lemma}
	\label{lem: Taylor_series}
	For any $0 \leq x \leq 1/2d$,
	\begin{equation*}
	(1 + x)^d \leq 1 + 2dx.
	\end{equation*}
\end{lemma}
\begin{proof}
	We take the binomial expansion of $(1 + x)^d$:
	\begin{align*}
	(1 + x)^d & = \sum_{k = 0}^{d} {d \choose k} x^k \\
	& = 1 + dx + dx\left(\sum_{k = 2}^d \frac{{d \choose k} x^{k - 1}}{d}\right)\\
	& \leq 1 + dx + dx\left(\sum_{k = 2}^d \frac{d^{k}}{(2d)^{k-1}d}\right) \\
	& \leq 1 + 2dx.
	\end{align*}
\end{proof}

We will repeatedly employ Lemma \ref{lem: expansion_sets} and Lemma \ref{lem: Taylor_series} in tandem. As a first example, in Lemma \ref{lem: interior_of_expansion_sets}, we use it to bound the ratio of $\nu(\Asig)$ to $\nu(\Aset_{\sigma - r})$. This will be useful when we bound $\vol(\Csig)$.

@@

\begin{lemma}
	\label{lem: expected_number_boundary_points}
	Under the conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $r < \sigma/2d$,
	\begin{equation*}
	\Pbb(\Csigr) \leq 2 \nu(\Csig) \frac{rd}{\sigma}  \left(\lambda_{\sigma} - \frac{r^{\gamma}}{\gamma + 1}\right)
	\end{equation*}	
\end{lemma}
\begin{proof}
	Recalling that $f$ is the density function for $\Pbb$, we have
	\begin{equation}
	\label{eqn: integral_over_epsilon_neighborhood}
	\Pbb(\Csigr) = \int_{\Csigr} f(x) dx
	\end{equation}
	We partition $\Csigr$ into slices, based on distance from $\Csig$, as follows: for $k \in \N$,
	\begin{equation*}
	\mathcal{T}_{i,k} = \set{x \in \Rd: t_{i,k} < \frac{\dist(x, \Csig)}{r} \leq t_{i+1,k}}, ~~ \Csigr = \bigcup_{i = 0}^{k-1} \mathcal{T}_{i,k}
	\end{equation*}
	where $t_i = i/k$ for $i = 0, \ldots, k - 1$. As a result,
	\begin{equation*}
	\int_{\Csigr} f(x) dx = \sum_{i = 0}^{k-1} \int_{\mathcal{T}_{i,k}} f(x) dx \leq \sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x).
	\end{equation*}
	We substitute
	\begin{equation*}
	\nu(\mathcal{T}_{i,k}) = \nu(\Csig + rt_{i+1,k}B) - \nu(\Csig + rt_{i,k}B) := \nu_{i+1,k} - \nu_{i,k}. 
	\end{equation*}
	where for simplicity we've written $\nu_{i,k} = \nu(\Csig + rt_{i+1,k}B)$.
	This, in concert with the upper bound
	\begin{equation*}
	\max_{x \in \mathcal{T}_{i,k}} f(x) \leq \lambda_{\sigma} - (rt_{i,k})^{\gamma},
	\end{equation*}
	which follows from \ref{asmp: bounded_density} and \ref{asmp: low_noise_density}, yields
	\begin{align}
	\label{eqn: telescoping_sum}
	\sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x) & \leq \sum_{i = 0}^{k-1} \biggl\{ \nu_{i+1,k} - \nu_{i,k} \biggr\} \biggl( \lambda_{\sigma} - (rt_{i,k})^{\gamma} \biggr) \nonumber \\
	& = \sum_{i = 1}^{k} 
	\underbrace{\nu_{i,k} \biggl( \left[\lambda_{\sigma} - (rt_{i,k})^{\gamma}\right] -  \left[\lambda_{\sigma} - (rt_{i-1,k})^{\gamma}\right]\biggr)}_{:= \Sigma_k} + \underbrace{\biggl(\nu_{k,k}\left[\lambda_{\sigma} - r^{\gamma}\right] - \nu_{1,k}\lambda_{\sigma} \biggr)}_{:= \xi_k}
	\end{align}

>>>>

\documentclass{article}
\renewcommand{\thesection}{\Alph{section}}
\renewcommand{\theequation}{A.\arabic{equation}}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{graphicx}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
% \usepackage{hyperref}
\usepackage{bm}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{xr-hyper}
\usepackage{natbib}
\usepackage{fullpage}

\let\pprspace\relax
\externaldocument{arxiv_local_density_clustering}

@@

Let $\Aset \subseteq \Reals^d$, and for $\sigma \geq 0$, write $\sigma B := B(0,\sigma) = \set{x \in \Rd: \norm{x} \leq \sigma}$ for the closed ball of radius $\sigma$ centered at the origin (and let $B^{\circ}(0,\sigma)$ denote the corresponding open ball). Let $\Asig = \Aset + \sigma B$ be the direct sum of $\Aset$ and $\sigma B$, $\Asig = \set{z = x + y: x \in \Aset, y \in \sigma B}$.

\begin{lemma}
	\label{lem: expansion_volume}
	If $\Aset$ is closed and bounded, then for any $\delta > 0$,
	\begin{equation*}
	\nu(\Asig + \delta B) \leq \left(1 + \frac{\delta}{\sigma}\right)^d \nu(\Asig).
	\end{equation*}
\end{lemma}
\begin{proof}
	We will show that for any $\epsilon > 0$, 
	\begin{equation}
	\label{eqn: ratio_of_volume}
	\frac{\nu(\Asig + \delta B)}{\nu(\Asig)} \leq \frac{(\sigma + \delta + \epsilon)^d}{\sigma^d}
	\end{equation}
	which is sufficient to prove the claim.

	Fix $\epsilon > 0$. Our first goal is to find a finite collection $x_1, \ldots, x_N \in \Rd$ such that
	\begin{equation*}
	\bigcup_{i = 1}^{N} B(x_i, \sigma) \subseteq \Asig \subset \bigcup_{i = 1}^{N} B(x_i, \sigma + \epsilon). \tag{$N := N(\epsilon)$}
	\end{equation*}

	Observe that since $\Aset$ is closed and bounded, it is compact. As $B(x,\sigma)$ is compact, and the direct sum of two compact sets is itself compact, $\Asig$ is compact. Moreover,
	\begin{equation*}
	\Asig \subset \bigcup_{x \in \Aset} B^{\circ}(x,\sigma + \epsilon)
	\end{equation*}
	so by compactness there exists $x_1, \ldots,x_N \in \Aset$ such that
	\begin{equation*}
	\Asig \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon).
	\end{equation*}

	By the triangle inequality, $\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \epsilon + \delta)$. Of course, for each $x_i \in \Aset$, $B(x_i,\sigma) \in \Asig$. Summarizing our findings, we have
	\begin{equation}
	\label{eqn: finite_subcover}
	\bigcup_{i = 1}^{N} B(x_i,\sigma) \subseteq \Asig  ,~\Asig + \delta B \subset \bigcup_{i = 1}^{N} B^{\circ}(x_i,\sigma + \delta + \epsilon)
	\end{equation}

	We next show a lower bound on $\nu(\Asig)$. Partition $\Asig$ using the balls $B(x_i,\sigma)$, meaning let $\Aset_{\sigma}^{(1)} := B(x_1,\sigma)$, $\Aset_{\sigma}^{(2)} := B(x_2,\sigma) \setminus B(x_1,\sigma)$, and continuing, so that
	\begin{equation*}
	\Aset_{\sigma}^{(i)} := B(x_i,\sigma) \setminus \bigcup_{j = 1}^{i - 1} \Aset_{\sigma}^{(j)}. \tag{$i = 1,\ldots,N$}
	\end{equation*}
	Observe that $\bigcup_{i = 1}^{N} \Asig^{(i)} = \bigcup_{i = 1}^{N} B(x_i,\sigma)$, so by \eqref{eqn: finite_subcover} $\Asig \supseteq \bigcup_{i = 1}^{N} \Asig^{(i)}$. As $\Asig^{(1)},\ldots, \Asig^{(N)}$ are non-overlapping,
	\begin{align*}
	\nu(\Asig) & \geq \sum_{i = 1}^{N} \nu(\Asig^{(i)}) \\
	& = \sigma^d \nu_d \sum_{i = 1}^{N}  \frac{\nu(\Asig^{(i)})}{\nu(B(x_i,\sigma))}
	\end{align*}
	We turn to proving an upper bound on $\nu(\Asig + \delta B)$. Let $\Aset_{\sigma + \epsilon + \delta}^{(1)} := B(x_1,\sigma + \delta + \epsilon)$ and
	\begin{equation*}
	\Aset_{\sigma + \delta + \epsilon}^{(i)} := B(x_i,\sigma + \delta + \epsilon) \setminus \bigcup_{j = 1}^{i - 1} \Aset_{\sigma + \delta + \epsilon}^{(j)}. \tag{$i = 2,\ldots,N$}
	\end{equation*}

	As $\bigcup_{i = 1}^{N} \Aset_{\sigma + \delta + \epsilon}^{(i)} = \bigcup_{i = 1}^{N} B(x_i,\sigma + \delta + \epsilon)$, by \eqref{eqn: finite_subcover}
	\begin{equation*}
	\Aset_{\sigma} + \delta B \subset \bigcup_{i =1}^{N} \Aset_{\sigma + \delta + \epsilon}^{(i)}
	\end{equation*}
	and therefore
	\begin{align*}
	\nu(\Aset_{\sigma + \delta}) & \leq \sum_{i = 1}^{N} \nu\bigl(\Aset_{\sigma + \delta + \epsilon}^{(i)}\bigr) \\
	& = \sum_{i = 1}^{N} \nu_d (\sigma + \delta + \epsilon)^d \frac{\nu(\Aset_{\sigma + \delta + \epsilon}^{(i)})}{\nu(B(x_i, \sigma + \delta + \epsilon))} \\
	& \leq \nu_d (\sigma + \delta + \epsilon)^d \sum_{i = 1}^{N} \frac{\nu(\Asig^{(i)})}{\nu(B(x_i,\sigma))}
	\end{align*}
	where the last inequality follows from Lemma \ref{lem: covering}. We have shown \eqref{eqn: ratio_of_volume}, and thus the claim.
\end{proof}

\begin{lemma}
	\label{lem: covering}
	For $i = 1, \ldots, N$ and  $A_{\sigma}^{(i)}, A_{\sigma + \delta + \epsilon}^{(i)}$ as in Theorem \ref{lem: expansion_volume},
	\begin{equation*}
	\frac{\nu(\Aset_{\sigma + \delta + \epsilon}^{(i)})}{\nu(B(x_i, \sigma + \delta + \epsilon))} \leq \frac{\nu(\Aset_{\sigma}^{(i)})}{\nu(B(x_i, \sigma))}
	\end{equation*}
\end{lemma}
\begin{proof}
	Let $\delta' := \delta + \epsilon$. It will be sufficient to show that
	\begin{equation*}
	\biggl(\Aset_{\sigma + \delta'}^{(i)} - \set{x_i}\biggr) \subseteq \left(1 + \frac{\delta'}{\sigma}\right)\cdot\biggl(\Asig^{(i)} - \set{x_i}\biggr) 
	\end{equation*}
	since then
	\begin{equation*}
	\nu(\Aset_{\sigma + \delta'}^{(i)}) \leq \left(1 + \frac{\delta'}{\sigma}\right)^d \nu(\Aset_{\sigma}^{(i)}) = \frac{\nu(B(x_i, \sigma + \delta'))}{\nu(B(x_i, \sigma))} \nu(\Aset_{\sigma}^{(i)}).
	\end{equation*}

	Assume without loss of generality that $x_i = 0$, and let $x \in \Aset_{\sigma + \delta'}^{(i)}$, meaning
	\begin{equation}
	\norm{x} \leq \sigma + \delta',~ \norm{x - x_j} > \sigma + \delta'~ \textrm{for $j = 1, \ldots, i - 1$}.
	\end{equation}
	Letting $x' = \frac{\sigma}{\sigma + \delta'} x$, since $\norm{x} \leq \sigma + \delta'$, $\norm{x'} \leq \sigma$ and therefore $x' \in B(0,\sigma)$. Additionally observe that for any $j = 1, \ldots, i - 1$, by the triangle inequality
	\begin{equation*}
	\norm{x' - x_j} \geq \norm{x - x_j} - \norm{x - x'} > \sigma + \delta' - \frac{\delta'}{\sigma + \delta'}\norm{x} \geq \sigma
	\end{equation*}
	and therefore $x' \not\in B(x_j,\sigma)$ for any $j = 1,\ldots, i - 1$. So $x' \in \Asig^{(i)}$.
\end{proof}

We will need to carefully control the volume of expansion sets using the estimate in Lemma \ref{lem: expansion_volume}; Lemma \ref{lem: Taylor_series} serves this purpose.
\begin{lemma}
	\label{lem: Taylor_series}
	For any $0 \leq x \leq 1/2d$,
	\begin{equation*}
	(1 + x)^d \leq 1 + 2dx.
	\end{equation*}
\end{lemma}
\begin{proof}
	We take the binomial expansion of $(1 + x)^d$:
	\begin{align*}
	(1 + x)^d & = \sum_{k = 0}^{d} {d \choose k} x^k \\
	& = 1 + dx + dx\left(\sum_{k = 2}^d \frac{{d \choose k} x^{k - 1}}{d}\right)\\
	& \leq 1 + dx + dx\left(\sum_{k = 2}^d \frac{{d \choose k}}{(2d)^{k-1}d}\right) \tag{since $x \leq \frac{1}{2d}$} \\
	& \leq 1 + dx + dx\left(\sum_{k = 2}^d \frac{1}{2^{k - 1}}\right) \leq 1 + 2dx.
	\end{align*}
\end{proof}

Let $\Csigr := \set{x: 0 < \dist(x, \Csig) < r}$, where $\Csig$ is as in Theorem \ref{thm: conductance_upper_bound}. Lemma \ref{lem: expected_number_boundary_points} involves the bulk of the technical effort required to prove Theorem \ref{thm: conductance_upper_bound}; it will be necessary to bound the expected cut size of $\Csig[\Xbf]$ in $G_{n,r}$.

\begin{lemma}
	\label{lem: expected_number_boundary_points}
	Under the conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{equation*}
	\Pbb(\Csigr) \leq \frac{2dr}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig)
	\end{equation*}	
\end{lemma}
\begin{proof}
	We partition $\Csigr$ into slices based on distance from $\Csig$ as follows: for $k \in \N$,
	\begin{equation*}
	\mathcal{T}_{i,k} = \set{x \in \Csigr: t_{i,k} < \frac{\dist(x, \Csig)}{r} \leq t_{i+1,k}}, ~~ \Csigr = \bigcup_{i = 0}^{k-1} \mathcal{T}_{i,k}
	\end{equation*}
	where $t_i = i/k$ for $i = 0, \ldots, k - 1$. As a result, for any $k \in \mathbb{N}$,
	\begin{equation}
	\label{eqn: partition_ub}
	\Pbb(\Csigr) = \int_{\Csigr} f(x) dx = \sum_{i = 0}^{k-1} \int_{\mathcal{T}_{i,k}} f(x) dx \leq \sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x).
	\end{equation}

	\ref{asmp: bounded_density} and \ref{asmp: low_noise_density} imply the upper bound
	\begin{equation*}
	\max_{x \in \mathcal{T}_{i,k}} f(x) \leq \lambda_{\sigma} - c_0(rt_{i,k})^{\gamma},
	\end{equation*}
	and writing
	\begin{equation*}
	\nu(\mathcal{T}_{i,k}) = \nu(\Csig + rt_{i+1,k}B) - \nu(\Csig + rt_{i,k}B) =: \nu_{i+1,k} - \nu_{i,k},
	\end{equation*}
	we have
	\begin{align}
	\label{eqn: telescoping_sum}
	\sum_{i = 0}^{k-1} \nu(\mathcal{T}_{i,k}) \max_{x \in \mathcal{T}_{i,k}} f(x) & \leq \sum_{i = 0}^{k-1} \biggl\{ \nu_{i+1,k} - \nu_{i,k} \biggr\} \biggl( \lambda_{\sigma} - c_0(rt_{i,k})^{\gamma} \biggr) \nonumber \\
	& = \sum_{i = 1}^{k} 
	\underbrace{\nu_{i,k} \biggl( \left[\lambda_{\sigma} - c_0(rt_{i-1,k})^{\gamma}\right] -  \left[\lambda_{\sigma} - c_0(rt_{i,k})^{\gamma}\right]\biggr)}_{:= \Sigma_k} + \underbrace{\biggl(\nu_{k,k}\left[\lambda_{\sigma} - c_0r^{\gamma}\right] - \nu_{1,k}\lambda_{\sigma} \biggr)}_{:= \xi}
	\end{align}
	where the second equality comes from rearranging terms in the sum.

	We first consider the term $\Sigma_k$. $\Cset$ has finite diameter by \ref{asmp: cluster_diameter}, moreover $\Cset$ is closed, due to the continuity of $f$. We may therefore apply Lemma \ref{lem: expansion_volume} to $\Csig$ and obtain
	\begin{equation*}
	\nu_{i,k} = \nu(\Csig + rt_{i,k}B) \leq \nu(\Csig)\left(1 + \frac{rt_{i,k}}{\sigma}\right)^d
	\end{equation*}
	which in turn gives
	\begin{align}
	\Sigma_k & \leq c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ rt_{i,k}}{\sigma}\right)^d \biggl( (t_{i,k})^{\gamma} - (t_{i-1,k})^{\gamma}\biggr) \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}).~~~~~~~~~~~~~~ (u_{i,k} = t_{i,k}^{\gamma}) \label{eqn: Sigmak_riemann_sum}
	\end{align}
	\eqref{eqn: Sigmak_riemann_sum} is a Riemann sum, and taking the limit as $k \to \infty$ we obtain
	\begin{align}
	\lim_{k \to \infty} c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}) & = c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{r u^{1/\gamma}}{\sigma}\right)^{d} du \nonumber \\
	& \overset{(i)}{\leq} c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{2 d r u^{1/\gamma}}{\sigma}\right) du \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \left(1 + \gamma \frac{2 d r}{(\gamma + 1)\sigma}\right). \label{eqn: Sigmak_integral}
	\end{align}
	where $(i)$ follows from Lemma \ref{lem: Taylor_series} in light of the fact $r \leq \sigma / 2d$.

	An upper bound on $\xi$ follows from largely the same logic, although it does not involve integration:
	\begin{align}
	\xi & \overset{(ii)}{\leq} \nu(\Csig) \biggl\{ \left(1 + \frac{ r}{\sigma}\right)^d(\lambda_{\sigma} - c_0r^{\gamma}) - \lambda_{\sigma} \biggr\} \nonumber \\
	& \overset{(iii)}{\leq} \nu(\Csig) \biggl\{ \left(1 + \frac{2d r}{\sigma}\right)(\lambda_{\sigma} - c_0r^{\gamma}) - \lambda_{\sigma} \biggr\} = \nu(\Csig) \biggl\{ \frac{2dr}{\sigma}(\lambda_{\sigma} - c_0r^{\gamma}) - c_0 r^{\gamma} \biggr\}. \label{eqn: xi_ub}
	\end{align}
	where $(ii)$ follows application of Lemma \ref{lem: expansion_volume} to bound $\nu_{k,k}$ and $(iii)$ from Lemma \ref{lem: Taylor_series}. As the bounds in \eqref{eqn: partition_ub} and \eqref{eqn: telescoping_sum} hold for all $k$, these along with \eqref{eqn: Sigmak_integral} and \eqref{eqn: xi_ub} imply the desired result.
\end{proof}

Lemma \ref{lem: local_conductance} will be necessary to lower bound the expected volume of $\Csig[\Xbf]$ in $G_{n,r}$. Define the \emph{local conductance} $\ell_{\nu,r}(u)$ to be
\begin{equation*}
\ell_{\nu,r}(u) = \frac{\nu\left(\Csig \cap B(u,r)\right)}{\nu\left(B(u,r)\right)}.
\end{equation*}
\begin{lemma}
	\label{lem: local_conductance}
	Let $u \in \Csig$. Then, for any $0 < r \leq \frac{\sigma}{2\sqrt{d}}$,
	\begin{equation*}
	\ell_{\nu,r}(u) \geq \frac{6}{25}. 
	\end{equation*}
\end{lemma}
\begin{proof}
	Since $u \in \Csig$ there exists $x \in \Cset$ such that $u \in B(x, \sigma)$, and as $B(x,\sigma) \subseteq \Csig$,
	\begin{equation*}
	\nu\bigl(B(u, r) \cap B(x, \sigma)\bigr) \leq \nu\bigl(B(u, r) \cap \Csig \bigr).
	\end{equation*}
	Without loss of generality, let $\norm{u - x} = \sigma$; it is not hard to see that if $\norm{u - x} < \sigma$, the volume of the overlap will only grow. Then, since $\norm{u  - x} = \sigma$, $B(u, r) \cap B(x, \sigma)$ contains a spherical cap of radius $r$ and height
	\begin{equation*}
	h = r - (r)^2/2\sigma = r \left( 1 - \frac{r}{2 \sigma} \right)
	\end{equation*}	
	which by Lemma \ref{lem: volume_of_spherical_cap} has volume
	\begin{equation*}
	\nu_{cap} = \frac{1}{2} \nu_d r^d I_{1 - \alpha}\left( \frac{d + 1}{2}  ,\frac{1}{2}\right)
	\end{equation*}
	with $\alpha = 1 - \frac{2rh - h^2}{r^2} = \frac{r^2}{4 \sigma^2} \leq \frac{1}{8d}$.

	Then by Lemmas \ref{lem: beta_integral} (applied with $t = 1$) and \ref{lem: beta_function},
	\begin{align*}
	I_{1 - \alpha}\left( \frac{d + 1}{2}  ,\frac{1}{2}\right) & \geq 1 - \frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \frac{3}{2\sqrt{d}} \\
	& \geq 1 - \frac{3}{4}\sqrt{\frac{d+2}{\pi d}} \geq 1 - \frac{3}{4}\sqrt{\frac{3}{2 \pi}}.
	\end{align*}
\end{proof}

The following formula for the volume of the spherical cap, stated in terms of the incomplete beta function, is well known. We include it without proof.

\begin{lemma}
	\label{lem: volume_of_spherical_cap}
	Let $\mathrm{Cap}_r(h)$ denote a spherical cap of radius $r$ and height $h$. Then, 
	\begin{equation*}
	\nu\bigl( \mathrm{Cap}_r(h)  \bigr) = \frac{1}{2} \nu_d r^d I_{1 - \alpha}(\frac{d + 1}{2}; \frac{1}{2})
	\end{equation*}
	where
	\begin{equation*}
	\alpha := 1 - \frac{2 r h - h^2}{r^2}
	\end{equation*}
	and
	\begin{equation*}
	I_{1 - \alpha}(z,w) = \frac{\Gamma(z + w)}{\Gamma(z) \Gamma(w)} \int_{0}^{1 - \alpha} u^{z - 1} (1 - u)^{w - 1} du.
	\end{equation*}
	is the cumulative distribution function of a $\mathrm{Beta}(z,w)$ distribution, evaluated at $1 - \alpha$. 
\end{lemma}

\begin{lemma}
	\label{lem: beta_integral}
	For any $0 \leq t \leq 1$ and $\alpha \leq \frac{t^2}{8 d}$,
	\begin{equation*}
	\int_{0}^{1 - \alpha}u^{(d-1)/2}(1 - u)^{-1/2}du \geq \frac{\Gamma\bigl(\frac{d + 1}{2}\bigr)\Gamma\bigl(\frac{1}{2}\bigr)}{ \Gamma\bigl(\frac{d}{2}+ 1\bigr)} - \frac{3t}{2\sqrt{d}}
	\end{equation*}
\end{lemma}
\begin{proof}
	We can write 
	\begin{equation*}
	\int_{0}^{1 - \alpha}u^{(d-1)/2}(1 - u)^{-1/2}du = \int_{0}^{1}u^{(d-1)/2}(1 - u)^{-1/2}du - \int_{1 - \alpha}^{1}u^{(d-1)/2}(1 - u)^{-1/2}du
	\end{equation*}
	The first integral is simply the beta function, with
	\begin{equation*}
	B(\frac{d+1}{2},\frac{1}{2}) := \frac{\Gamma\bigl(\frac{d + 1}{2}\bigr)\Gamma\bigl(\frac{1}{2}\bigr)}{ \Gamma\bigl(\frac{d}{2}+ 1\bigr)}.
	\end{equation*}
	To upper bound the second integral, we apply the Taylor theorem with remainder to $(1 - u)^{-1/2}$, obtaining
	\begin{equation*}
	(1 - u)^{-1/2} \leq \alpha^{-1/2} + \max_{u \in (1 - \alpha, 1)} \frac{\alpha}{2} (1 - u)^{-3/2} = \frac{3}{2}\alpha^{-1/2}.
	\end{equation*}
	As a result,
	\begin{align*}
	\int_{1 - \alpha}^{1}u^{(d-1)/2}(1 - u)^{-1/2}du & \leq \frac{3}{2}\alpha^{-1/2} \int_{1 - \alpha}^{1}u^{(d-1)/2}du \\
	& = \frac{3}{d+1}\alpha^{-1/2} \left(1 - (1 - \alpha)^{(d + 1)/2}\right) \\
	& \overset{(iv)}{\leq} \frac{3}{(d+1)}\alpha^{-1/2} (\alpha(d + 1)) \\
	& =  3\alpha^{1/2}.
	\end{align*}
	where $(iv)$ follows from Lemma \ref{lem: Taylor_series}, and the condition $\alpha \leq \frac{t^2}{8d}$. The result follows from the condition $\alpha \leq \frac{t^2}{8d}$. 
\end{proof}

Lemma \ref{lem: beta_function} follows from $\Gamma(1/2) = \sqrt{\pi}$ and the upper bound $\Gamma(x + 1)/ \Gamma(x+s) \leq (x + 1)^{1-s}$ for $s \in [0,1]$.
\begin{lemma}
	\label{lem: beta_function}
	\begin{equation*}
	\frac{\Gamma\bigl(\frac{d}{2}+ 1\bigr)}{\Gamma\bigl(\frac{d + 1}{2}\bigr) \Gamma\bigl(\frac{1}{2}\bigr)} \leq \sqrt{\frac{d + 2}{2\pi}}
	\end{equation*}
\end{lemma}

@@

For notational ease, we write
\begin{align*}
\cut_{n,r} = \cut(\Csig[\Xbf]; G_{n,r}), ~ \mu_K = \mathbb{E}(\cut_{n,r}), ~ p_K = \frac{\mu_K}{{n \choose 2}} \\
\vol_{n,r} = \vol(\Csig[\Xbf]; G_{n,r}), ~ \mu_V = \mathbb{E}(\vol_{n,r}), ~ p_V = \frac{\mu_V}{{n \choose 2}} \\
\vol_{n,r}^c = \vol(\Xbf \setminus \Csig[\Xbf]; G_{n,r}), ~ \mu_V^c = \mathbb{E}(\vol_{n,r}^c), ~ p_V^c = \frac{\mu_V^c}{{n \choose 2}}
\end{align*}
for the random variable, mean, and probability of cut size and volume, respectively.

\begin{lemma}
	\label{lem: expected_density_cut}
	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{equation*}
	p_K \leq \frac{4 d \nu_d r^{d+1} \lambda}{\sigma} \left(\lambda_{\sigma} - c_0\frac{r^{\gamma}}{\gamma + 1}\right) \nu(\Csig)
	\end{equation*}
\end{lemma}
\begin{proof}
	We can write $\cut_{n,r}$ as a double sum,
	\begin{equation}
	\label{eqn: density_cut_expansion}
	\cut_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \not\in \Csig) \1(x_j \in \Csig) \1(\norm{x_i - x_j} \leq r)
	\end{equation}
	and by linearity of expectation, we obtain
	\begin{equation*}
	p_K = \frac{\mu_K}{{n \choose 2}} = 2 \cdot \Pbb(x_i \not\in \Csig, x_j \in \Csig, \norm{x_i - x_j} \leq r). \tag{for each $i,j$, $i \neq j$}
	\end{equation*}
	Writing this with respect to the density function $f$, we have
	\begin{align*}
	p_K & = 2 \int_{\Rd \setminus \Csig} f(x) \Pbb\bigl(B(x,r) \cap \Csig\bigr) dx \\
	& = 2 \int_{\Csigr} f(x) \Pbb\bigl(B(x,r) \cap \Csig\bigr) dx \\
	& \leq 2 \nu_d r^d \lambda  \int_{\Csigr} f(x) dx = 2 \nu_d r^d \lambda \Pbb(\Csigr).
	\end{align*}
	where the inequality follows from \ref{asmp: cluster_separation}, which implies $f(x) \leq \lambda$ for $x \in \Csig \setminus \Cset$. Then, upper bounding the integral using Lemma \ref{lem: expected_density_cut} gives the final result.
\end{proof}

\begin{lemma}
	\label{lem: expected_density_volume}
	Under the setup and conditions of Theorem \ref{thm: conductance_upper_bound}, and for any $0 < r \leq \sigma/2d$,
	\begin{equation*}
	p_V \geq \frac{12}{25} \lambda_{\sigma}^2 \nu_d r^d \nu(\Csig)
	\end{equation*}
\end{lemma}
\begin{proof}
	The proof will proceed similarly to Lemma \ref{lem: expected_density_cut}. We begin by writing $\vol_{n,r}$ as the sum of indicator functions,
	\begin{equation}
	\label{eqn: volume_expansion}
	\vol_{n,r} = \sum_{i = 1}^{n} \sum_{j \neq i} \1(x_i \in \Csig) \1(x_j \in B(x_i, r))
	\end{equation}
	and by linearity of expectation we obtain
	\begin{equation*}
	p_V = \frac{\mu_V}{{n \choose 2}} = 2 \cdot \Pbb(x_i \in \Csig, x_j \in B(x_i,r)). \tag{for any $i,j$, $i \neq j$. }
	\end{equation*}
	Writing this with respect to the density function $f$, we have
	\begin{align*}
	p_V & = 2 \int_{\Csig} f(x) \Pbb(B(x,r)) dx \\
	& \geq 2 \int_{\Csig} f(x) \Pbb(B(x,r) \cap \Csig) dx
	\end{align*}
	whence the claim then follows by Lemma \ref{lem: local_conductance}.
\end{proof}

To employ Lemmas \ref{lem: expected_density_cut} and \ref{lem: expected_density_volume} in the proof of Theorem \ref{thm: conductance_upper_bound}, we must relate the random variable
\begin{equation*}
\Phi_{n,r}(\Csig[\Xbf]) = \frac{\cut_{n,r}}{\min \set{\vol_{n,r}, \vol_{n,r}^c}}
\end{equation*}
to $p_K$ and $p_V$.
